# Phenotypes

```{r}
library(tidyverse)
# Get lighter/darker functions
source("https://gist.githubusercontent.com/brettellebi/c5015ee666cdf8d9f7e25fa3c8063c99/raw/91e601f82da6c614b4983d8afc4ef399fa58ed4b/karyoploteR_lighter_darker.R")
```

## Read in phenotype data

```{r}
pheno_file = here::here("data/20220214_phenotypes.xlsx")

df = readxl::read_xlsx(pheno_file) %>% 
  # remove "KC" from sample name
  dplyr::mutate(sample = fish %>% 
                  stringr::str_remove("KC"))

# How many samples?
length(unique(df$sample))
```

## Period 

Note that all F0s and F1s were imaged on the AU scope. 

DB tends to have a higher temperature than AU, by 0.7°C. 

AU has been extensively calibrated, which may cause the differences in the phenotype measurements.

### Mean

Difference between means of two microscopes:

```{r}
micr_mean = df %>% 
  dplyr::filter(!is.na(Microscope)) %>% 
  dplyr::group_by(Microscope) %>% 
  dplyr::summarise(mean_pheno = mean(mean))

micr_mean

# Difference
diff_mean = abs(micr_mean$mean_pheno[1] - micr_mean$mean_pheno[2])
diff_mean
```

#### Historgram

```{r}
mean_pal = c("#177e89", "#084c61")

df %>% 
  # remove NAs in `Microscope` column
  dplyr::filter(!is.na(Microscope)) %>% 
  dplyr::group_by(Microscope) %>% 
  ggplot() +
    geom_histogram(aes(x = mean, fill = Microscope),
                   bins = 50) +
    scale_fill_manual(values = mean_pal) +
    geom_vline(data=micr_mean, aes(xintercept=mean_pheno), colour = darker(mean_pal, amount = 50)) +
    #stat_summary(fun=mean, geom="line") +
    facet_grid(rows = vars(Microscope)) +
    theme_bw() +
    xlab("period mean") +
    ggtitle(paste("Period mean\n", "Difference between means: ", diff_mean, sep = ""))
  
```

#### Q-Q plot

```{r}
df %>% 
  # remove NAs in `Microscope` column
  dplyr::filter(!is.na(Microscope)) %>% 
  ggplot() +
    stat_qq(aes(sample = mean, colour = Microscope)) +
    stat_qq_line(aes(sample = mean, colour = Microscope)) +
    scale_colour_manual(values = c("#177e89", "#084c61")) +
    facet_grid(rows = vars(Microscope)) +
    theme_bw() +
    xlab("period mean") +
    ggtitle("Period mean") +
    theme(aspect.ratio=1)
  
```

## Intercept

Difference between means of two microscopes:

```{r}
micr_intercept = df %>% 
  dplyr::filter(!is.na(Microscope)) %>% 
  dplyr::group_by(Microscope) %>% 
  dplyr::summarise(mean_pheno = mean(intercept))

micr_intercept

# Difference
diff_intercept = abs(micr_intercept$mean_pheno[1] - micr_intercept$mean_pheno[2])
diff_intercept
```

#### Histogram 

```{r}
intercept_pal = c("#8D99AE", "#2b2d42")

df %>% 
  # remove NAs in `Microscope` column
  dplyr::filter(!is.na(Microscope)) %>% 
  ggplot() +
    geom_histogram(aes(intercept, fill = Microscope),
                   bins = 50) +
    geom_vline(data=micr_intercept, aes(xintercept=mean_pheno), colour = darker(intercept_pal, amount = 50)) +
    scale_fill_manual(values = intercept_pal) +
    facet_grid(rows = vars(Microscope)) +
    xlab("period intercept") +
    theme_bw() +
    ggtitle(paste("Period intercept\n", "Difference between means: ", diff_intercept, sep = ""))
  
```

#### Q-Q plot

```{r}
df %>% 
  # remove NAs in `Microscope` column
  dplyr::filter(!is.na(Microscope)) %>% 
  ggplot() +
    stat_qq(aes(sample = intercept, colour = Microscope)) +
    stat_qq_line(aes(sample = intercept, colour = Microscope)) +
    scale_colour_manual(values = c("#8D99AE", "#2b2d42")) +
    facet_grid(rows = vars(Microscope)) +
    theme_bw() +
    xlab("period intercept") +
    ggtitle("Period intercept") +
    theme(aspect.ratio=1)
```


## Presomitic mesoderm size

### Histogram 

```{r}
df %>% 
  ggplot() +
    geom_histogram(aes(unsegmented_psm_area),
                   fill = "#401f3e",
                   bins = 50) +
    theme_bw() +
    ggtitle("Presomitic mesoderm size")
```

### Q-Q plot

```{r}
df %>% 
  ggplot(aes(sample = intercept)) +
    stat_qq(colour = "#401f3e") +
    stat_qq_line(colour = darker("#401f3e")) +
    theme_bw() +
    xlab("presomitic mesoderm area") +
    ggtitle("Presomitic mesoderm area") +
    theme(aspect.ratio=1)
```