# Experiment with adding covariates

```{r}
library(tidyverse)
library(GridLMM)
library(KRLS)
```

## No covariates

```{r}
covariates = NULL
```


### Read in data

```{r}
GENO_FILE = "/nfs/research/birney/users/ian/somites/association_testing/20220214/all_sites/inputs/5000.rds"
PHENO_FILE = here::here("data/20220214_phenotypes.xlsx") # True phenotypes
#GWLS_SOURCE_FILE = here::here("workflow/scripts/run_gwls_source.R")
MANHAT_SOURCE_FILE = here::here("book/source/04-Association_testing_with_covariates.R")
BIN_LENGTH = 5000
TARGET_PHENO = "unsegmented_psm_area"
COVARIATES = "none"
PLOT_DIR = here::here("book/plots/20220214/covariate_test")

# Read in source files
source(MANHAT_SOURCE_FILE)

invers_norm=F
```

#### Genotypes and positions

```{r}
# Load genotypes and positions
in_list = readRDS(GENO_FILE)
```

#### Phenotypes

```{r}
## Read in file and wrangle
phenos = readxl::read_xlsx(PHENO_FILE) %>%
    # adjust sample names
    dplyr::mutate(SAMPLE = fish %>% stringr::str_remove("KC")) %>%
    # select key columns
    dplyr::select(SAMPLE, all_of(TARGET_PHENO), Microscope, Date_of_imaging) %>%
    # ensure that the phenotype column is numeric
    dplyr::mutate(dplyr::across(all_of(TARGET_PHENO),
                                ~ as.numeric(.x))) 

```

#### Filter for samples with both genotypes and phenotypes

```{r}
## Filter and order phenotypes
in_list[["phenotypes"]] = phenos %>%
    # filter phenotypes for those with genotypes
    dplyr::filter(SAMPLE %in% in_list[["sample_order"]]) %>%
    # join to `sample_order` to ensure phenotypes are in the correct order   
    dplyr::left_join(tibble::tibble(SAMPLE = in_list[["sample_order"]]),
                     .,
                     by = "SAMPLE") %>%
    # remove NAs (created by the samples that have genotypes but not phenotypes)
    tidyr::drop_na() %>%
    # the GridLMM code doesn't work with tibbles
    as.data.frame()

## Filter genotypes for those that have phenotypes
in_list[["genotypes"]] = in_list[["genotypes"]] %>%
    dplyr::filter(in_list[["sample_order"]] %in% in_list[["phenotypes"]]$SAMPLE)

## Filter sample_order for those that have phenotypes
in_list[["sample_order"]] = in_list[["phenotypes"]]$SAMPLE

## Get number of samples
N_SAMPLES = nrow(in_list[["phenotypes"]])
```

### Run GWLS

#### Prepare data

```{r, eval = F}
# Rename key data
d = in_list[["genotypes"]]
m = in_list[["positions"]]
p = in_list[["phenotypes"]]

  # Creates a vector of the prefix "snp" combined with the row number of positions
  ids = paste("snp", 1:nrow(m), sep="")
  # Adds this vector as first column of the DF of positions: now comprising `ids`, `CHROM`, `BIN_START`, `BIN_END`
  mm = data.frame(ids, m)
  # Also add these ids as column names to the sample x locus DF of genotypes 
  colnames(d) = ids
  # Rename columns of positions DF
  colnames(mm) = c("snp", "Chr", "pos")
  # Then pull just the first three columns
  mm = mm[,1:3]
  # Rename it as `map`
  map = mm
  # Get number of samples (rows) from genotype DF
  n_geno=nrow(d)
  # Create matrix of 1200 elements from 1 to 1200, with 10 columns
  Plots = matrix(1:1200,nc = 10)
  # Create matrix of row numbers for each element of the `Plots` matrix (1:1200)
  Plot_Row = row(Plots)
  # Create matrix of col numbers for each element of the `Plots` matrix (1:10)
  Plot_Col = col(Plots)
  # Create a data frame with "Genox" for each sample in the first column
  # And randomly sample `n_geno` times from `Plots`
  data = data.frame(Geno = paste0('Geno',1: n_geno), Plot = sample(Plots)[1:nrow(d)])
  # Add a column `Row` with the row number of each randomly-sampled element
  data$Row = Plot_Row[data$Plot]
  # Add a column `Col` with the row number of each randomly-sampled element
  data$Col = Plot_Col[data$Plot]
  # Add phenotype
  if(invers_norm) {
    ## Inverse-normalise
    data = cbind(data,
                 my.invnorm(p %>% 
                              dplyr::select(y = dplyr::all_of(TARGET_PHENO))), # inverse-normalise phenotype
                 p %>% 
                   dplyr::select(dplyr::all_of(covariates))) # pull covariates
  } else {
    data = cbind(data,
                 p %>% 
                   dplyr::select(y = dplyr::all_of(TARGET_PHENO), # pull phenotype
                                 dplyr::all_of(covariates))) # and covariates
  }
  # Convert genotypes DF to matrix
  X = as.matrix(d)
  # Convert NAs to 0
  X[is.na(X)]=0
  # Add sample names to `X` as row names
  row.names(X) = data[,1]
  # Center genotypes by subtracting the mean for each marker (column) from the genotype value
  X_centered = sweep(X,2,colMeans(X),'-') # center marker genotypes
  # Use centered genotypes to create `K` kinship matrix
  K = tcrossprod(X_centered) / ncol(X_centered)
  # Make rownames and colnames of `K` the randomly sampled elements from `Plots`
  rownames(K) = colnames(K) = data$Plot
  # Calculate the spatial kernel
  field = data[,c('Row','Col')] # Pull out `Row` and `Col` columns
  dists = as.matrix(dist(field)) # Compute the distances 
  h = median(dists) # Calculate default tuning parameter `h`
  K_plot = KRLS::gausskernel(field,h^2/2); diag(K_plot)=1 # Compute the NxN distance matrix with pairwise with pairwise differences between rows as measured by a Gaussian Kernel
  rownames(K_plot) = colnames(K_plot) = data$Plot
  # Create `test_formula`
  if (is.null(covariates)){
    TEST_FORMULA = "~1"
  } else if (!is.null(covariates)){
    TEST_FORMULA = paste("~1 +", paste(covariates, collapse = " + "))
  }
  TEST_FORMULA = as.formula(TEST_FORMULA)
```

#### Run regular GWAS 

```{r, eval = F}
# Run gwas
gwas_original = GridLMM::GridLMM_GWAS(
  formula = y~1 + (1|Geno), # the same error model is used for each marker. It is specified similarly to lmer
  test_formula = TEST_FORMULA, # this is the model for each marker. ~1 means an intercept for the marker. ~1 + cov means an intercept plus a slope on `cov` for each marker
  reduced_formula = ~0, # This is the null model for each test. ~0 means just the error model. ~1 means the null includes the intercept of the marker, but not anything additional
  data = data, # The dataframe to look for terms from the 3 models
  weights = NULL, # optional observation-specific weights
  X = X, # The matrix of markers. Note: This can be of dimension n_g x p, where n_g is the number of unique genotypes.
  X_ID = 'Geno', # The column of the data that identifies each genotype. Each level of data$Geno should match a rowname of X
  h2_start = NULL, # An optional vector of h2s to use as starting values for each random effect in the model. If NULL, will be calculated from the error model using GridLMM_ML
  h2_step = 0.01, # step size per random effect for testing alternate values of h2
  max_steps = 100, # maximum number of steps of size h2_step to take from h2_start
  X_map = map, # Optional. The marker positions.
  relmat = list(Plot = K), # A list of Kernel matrices for the random effects. If X_ID (here Geno) is not included in this list, then it is calculated as tcrossprod(Xc)/ncol(Xc) where Xc is the centered (and optionally scaled) X. If any random effects are described in `error_model` but not provided here, the Kernel is assumed to be the identity matrix
  centerX = TRUE, # Should the markers be centered when calculating the GRM (only will be done if needed for calculating the GRM),
  scaleX = FALSE, # Should the markers be scaled to have constant variance when calculating the GRM?
  fillNAX = FALSE, # Should missing marker data be filled in with the mean allele frequency?
  method = 'REML', # REML = Wald test, ML = LRT, BF = calculate Bayes factors
  mc.cores = my_detectCores(), # How many cores should be used for parallel processing. Unless X is large, tends to actually be faster with mc.cores = 1
  verbose = FALSE # Should progress be printed to the screen?
)

saveRDS(gwas_original, paste("/hps/nobackup/birney/users/ian/somites/covariate_test/gwas_results/", COVARIATES, ".rds", sep = ""))
```

```{r}
gwas_original = readRDS(paste("/hps/nobackup/birney/users/ian/somites/covariate_test/gwas_results/", COVARIATES, ".rds", sep = ""))
```

### Run permutations

#### Permute phenos

```{r, eval = F}
seeds = 1:10

counter = 0
perm_phenos = purrr::map(seeds, function(SEED){
  counter <<- counter + 1
  # set seed
  set.seed(seeds[counter])
  # randomise
  phenos = phenos %>% 
      # randomise phenotype
      dplyr::mutate(dplyr::across(all_of(TARGET_PHENO),
                                  ~ sample(.x)))
})
```

#### Run GWLS

```{r, eval = F}
perm_out = purrr::map(perm_phenos, function(PERM_PHENO){
  
  ## Filter and order phenotypes
  in_list[["phenotypes"]] = PERM_PHENO %>%
      # filter phenotypes for those with genotypes
      dplyr::filter(SAMPLE %in% in_list[["sample_order"]]) %>%
      # join to `sample_order` to ensure phenotypes are in the correct order   
      dplyr::left_join(tibble::tibble(SAMPLE = in_list[["sample_order"]]),
                       .,
                       by = "SAMPLE") %>%
      # remove NAs (created by the samples that have genotypes but not phenotypes)
      tidyr::drop_na() %>%
      # the GridLMM code doesn't work with tibbles
      as.data.frame()
  
  ## Filter genotypes for those that have phenotypes
  in_list[["genotypes"]] = in_list[["genotypes"]] %>%
      dplyr::filter(in_list[["sample_order"]] %in% in_list[["phenotypes"]]$SAMPLE)
  
  ## Filter sample_order for those that have phenotypes
  in_list[["sample_order"]] = in_list[["phenotypes"]]$SAMPLE
              
  # Run GWAS
  
  out = run_gwas(d = in_list[["genotypes"]],
                 m = in_list[["positions"]],
                 p = in_list[["phenotypes"]]
                )
  
  return(out)
})

names(perm_out) = seeds

saveRDS(perm_out, paste("/hps/nobackup/birney/users/ian/somites/covariate_test/perms/", COVARIATES, ".rds", sep = ""))
```

```{r}
perm_out = readRDS(paste("/hps/nobackup/birney/users/ian/somites/covariate_test/perms/", COVARIATES, ".rds", sep = ""))
```

### Get minimum

```{r}
perm_df = purrr::map_dfr(perm_out, function(PERM){
  OUT = tibble::tibble(MIN_P = PERM$results$p_value_REML %>%
                         min(., na.rm = T)
  )
}, .id = "SEED")

# Get minimum
SIG_LEVEL = min(perm_df$MIN_P)
```

### Generate Manhattan plot

```{r}
out_clean = clean_gwas_res(gwas_original,
                           bin_length = BIN_LENGTH,
                           chr_lens = med_chr_lens)

# Plot
out_plot = plot_man(out_clean,
                    site_filter = "all_sites",
                    phenotype = TARGET_PHENO,
                    bin_length = BIN_LENGTH, 
                    gwas_pal = eval(as.name(paste(TARGET_PHENO, "_pal", sep = ""))),
                    med_chr_lens = med_chr_lens,
                    sig_level = SIG_LEVEL) +
                 ylim(0,7) + 
                 labs(subtitle = paste("No covariates\nn samples: ", N_SAMPLES, sep = ""))

out_plot

```

```{r}
ggsave(file.path(PLOT_DIR, paste(COVARIATES, ".png", sep = "")),
       out_plot,
       device = "png",
       width = 9.6,
       height = 6,
       units = "in",
       dpi = 400)
```


## With Microscope as a covariate

```{r}
COVARIATES = "Microscope"
```

### GWAS

```{r, eval = F}
gwas_microscope = run_gwas(d = in_list[["genotypes"]],
                           m = in_list[["positions"]],
                           p = in_list[["phenotypes"]],
                           covariates = COVARIATES)

saveRDS(gwas_microscope, paste("/hps/nobackup/birney/users/ian/somites/covariate_test/gwas_results/",
                               paste(COVARIATES, collapse = "-"),
                               ".rds",
                               sep = ""))
```

```{r}
gwas_microscope = readRDS(paste("/hps/nobackup/birney/users/ian/somites/covariate_test/gwas_results/",
                                paste(COVARIATES, collapse = "-"),
                                ".rds",
                                sep = ""))
```


### Permutations

```{r, eval = F}
perm_out_microscope = purrr::map(perm_phenos, function(PERM_PHENO){
  
  ## Filter and order phenotypes
  in_list[["phenotypes"]] = PERM_PHENO %>%
      # filter phenotypes for those with genotypes
      dplyr::filter(SAMPLE %in% in_list[["sample_order"]]) %>%
      # join to `sample_order` to ensure phenotypes are in the correct order   
      dplyr::left_join(tibble::tibble(SAMPLE = in_list[["sample_order"]]),
                       .,
                       by = "SAMPLE") %>%
      # remove NAs (created by the samples that have genotypes but not phenotypes)
      tidyr::drop_na() %>%
      # the GridLMM code doesn't work with tibbles
      as.data.frame()
  
  ## Filter genotypes for those that have phenotypes
  in_list[["genotypes"]] = in_list[["genotypes"]] %>%
      dplyr::filter(in_list[["sample_order"]] %in% in_list[["phenotypes"]]$SAMPLE)
  
  ## Filter sample_order for those that have phenotypes
  in_list[["sample_order"]] = in_list[["phenotypes"]]$SAMPLE
              
  # Run GWAS
  
  out = run_gwas_with_covariates(d = in_list[["genotypes"]],
                                 m = in_list[["positions"]],
                                 p = in_list[["phenotypes"]],
                                 covariates = COVARIATES)
  
  return(out)
})

names(perm_out_microscope) = seeds

saveRDS(perm_out_microscope, paste("/hps/nobackup/birney/users/ian/somites/covariate_test/perms/", paste(COVARIATES, collapse = "-"), ".rds", sep = ""))
```

```{r}
perm_out_microscope = readRDS(paste("/hps/nobackup/birney/users/ian/somites/covariate_test/perms/",
                              paste(COVARIATES, collapse = "-"),
                              ".rds",
                              sep = ""))
```

### Get minimum

```{r}
perm_df = purrr::map_dfr(perm_out_microscope, function(PERM){
  OUT = tibble::tibble(MIN_P = PERM$results$p_value_REML.1 %>%
                         min(., na.rm = T)
  )
}, .id = "SEED")

# Get minimum
SIG_LEVEL = min(perm_df$MIN_P)
```

### Generate Manhattan plot

```{r}
out_clean = clean_gwas_res(gwas_microscope,
                           bin_length = BIN_LENGTH,
                           chr_lens = med_chr_lens) %>% 
  # There are now 2 p_value_REML columns: p_value_REML.1 and p_value_REML.2
  dplyr::rename(p_value_REML = p_value_REML.1)

# Plot
out_plot = plot_man(out_clean,
                    site_filter = "all_sites",
                    phenotype = TARGET_PHENO,
                    bin_length = BIN_LENGTH, 
                    gwas_pal = eval(as.name(paste(TARGET_PHENO, "_pal", sep = ""))),
                    med_chr_lens = med_chr_lens,
                    sig_level = SIG_LEVEL) +
                 ylim(0,7) + 
                 labs(subtitle = paste("Covariates: ",
                                       paste(COVARIATES, collapse = " & "),
                                       "\nn samples: ",
                                       N_SAMPLES,
                                       sep = ""))

out_plot

```

## With Microscope and Date_of_imaging as covariates

```{r}
COVARIATES = c("Microscope", "Date_of_imaging")
```


### GWAS

```{r}
gwas_mic-tod = run_gwas_with_covariates(d = in_list[["genotypes"]],
                                        m = in_list[["positions"]],
                                        p = in_list[["phenotypes"]],
                                        covariates = COVARIATES)

saveRDS(gwas_mic-tod, paste("/hps/nobackup/birney/users/ian/somites/covariate_test/gwas_results/",
                               paste(COVARIATES, collapse = "-"),
                               ".rds",
                               sep = ""))
```

```{r}
gwas_mic-tod = readRDS(paste("/hps/nobackup/birney/users/ian/somites/covariate_test/gwas_results/",
                                paste(COVARIATES, collapse = "-"),
                                ".rds",
                                sep = ""))
```


### Permutations

```{r}
perm_out_mic-tod = purrr::map(perm_phenos, function(PERM_PHENO){
  
  ## Filter and order phenotypes
  in_list[["phenotypes"]] = PERM_PHENO %>%
      # filter phenotypes for those with genotypes
      dplyr::filter(SAMPLE %in% in_list[["sample_order"]]) %>%
      # join to `sample_order` to ensure phenotypes are in the correct order   
      dplyr::left_join(tibble::tibble(SAMPLE = in_list[["sample_order"]]),
                       .,
                       by = "SAMPLE") %>%
      # remove NAs (created by the samples that have genotypes but not phenotypes)
      tidyr::drop_na() %>%
      # the GridLMM code doesn't work with tibbles
      as.data.frame()
  
  ## Filter genotypes for those that have phenotypes
  in_list[["genotypes"]] = in_list[["genotypes"]] %>%
      dplyr::filter(in_list[["sample_order"]] %in% in_list[["phenotypes"]]$SAMPLE)
  
  ## Filter sample_order for those that have phenotypes
  in_list[["sample_order"]] = in_list[["phenotypes"]]$SAMPLE
              
  # Run GWAS
  
  out = run_gwas_with_covariates(d = in_list[["genotypes"]],
                                 m = in_list[["positions"]],
                                 p = in_list[["phenotypes"]],
                                 covariates = COVARIATES)
  
  return(out)
})

names(perm_out_mic-tod) = seeds

saveRDS(perm_out_mic-tod, paste("/hps/nobackup/birney/users/ian/somites/covariate_test/perms/", paste(COVARIATES, collapse = "-"), ".rds", sep = ""))
```

```{r}
perm_out_microscope = readRDS(paste("/hps/nobackup/birney/users/ian/somites/covariate_test/perms/",
                              paste(COVARIATES, collapse = "-"),
                              ".rds",
                              sep = ""))
```

### Get minimum

```{r}
perm_df = purrr::map_dfr(perm_out_microscope, function(PERM){
  OUT = tibble::tibble(MIN_P = PERM$results$p_value_REML.1 %>%
                         min(., na.rm = T)
  )
}, .id = "SEED")

# Get minimum
SIG_LEVEL = min(perm_df$MIN_P)
```

### Generate Manhattan plot

```{r}
out_clean = clean_gwas_res(gwas_mic-tod,
                           bin_length = BIN_LENGTH,
                           chr_lens = med_chr_lens) %>% 
  # There are now 2 p_value_REML columns: p_value_REML.1 and p_value_REML.2
  dplyr::rename(p_value_REML = p_value_REML.1)

# Plot
out_plot = plot_man(out_clean,
                    site_filter = "all_sites",
                    phenotype = TARGET_PHENO,
                    bin_length = BIN_LENGTH, 
                    gwas_pal = eval(as.name(paste(TARGET_PHENO, "_pal", sep = ""))),
                    med_chr_lens = med_chr_lens,
                    sig_level = SIG_LEVEL) +
                 ylim(0,7) + 
                 labs(subtitle = paste("Covariates: ",
                                       paste(COVARIATES, collapse = " & "),
                                       "\nn samples: ",
                                       N_SAMPLES,
                                       sep = ""))

out_plot

```